---
title: "How Does A Bike Share Navigate Speedy Success"
author: "Muhamad Amirul Adham"
date: "`r Sys.Date()`"
output: html_document
---

```{r install, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

install.packages("tidyverse", repos = "http://cran.us.r-project.org")
install.packages("janitor", repos = "http://cran.us.r-project.org")
install.packages("lubridate", repos = "http://cran.us.r-project.org")
install.packages("dplyr", repos = "http://cran.us.r-project.org")
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
library(tidyverse)
library(janitor)
library(lubridate)
library(dplyr)
library(ggplot2)
```


## Introduction

Cyclistic: A bike-share program that features more than 5,800 bicycles and 600 docking stations. Cyclistic sets itself apart by also offering reclining bikes, hand tricycles, and cargo bikes, making bike-share more inclusive to people with disabilities and riders who can’t use a standard two-wheeled bike.

The director of marketing believes the company’s future success depends on maximizing the number of annual memberships. Therefore, the team wants to understand how casual riders and annual members use Cyclistic bikes differently. From these insights, the team will design a new marketing strategy to convert casual riders into annual members.

## Objective

3 main questions need to be answered before proceeding to the next step.

How do annual members and casual riders use Cyclistic bikes differently?

Why would casual riders buy Cyclistic annual memberships?

How can Cyclistic use digital media to influence casual riders to become members?

## Data

The latest 12 month data will be collected from [Cyclistic historical trip data](https://divvy-tripdata.s3.amazonaws.com/index.html) for the analysis.

## Analysis 

### RStudio

Upload all the 12 separate csv file into RStudio and merge it as a one dataset

```{r data, echo=FALSE}

d1<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202212-divvy-tripdata.csv")
d2<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202301-divvy-tripdata.csv")
d3<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202302-divvy-tripdata.csv")
d4<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202303-divvy-tripdata.csv")
d5<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202304-divvy-tripdata.csv")
d6<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202305-divvy-tripdata.csv")
d7<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202306-divvy-tripdata.csv")
d8<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202307-divvy-tripdata.csv")
d9<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202308-divvy-tripdata.csv")
d10<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202309-divvy-tripdata.csv")
d11<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202310-divvy-tripdata.csv")
d12<-read.csv("D:/Tupa/Data Analyst/Case Study/How Does A Bike Share Navigate Speedy Success/202311-divvy-tripdata.csv")
trip <- rbind(d1, d2, d3, d4, d5, d6, d7, d8, d9, d10, d11, d12)
head(trip)
```

The final outcome of dataset after some cleaning and filtering

```{r cleaning, echo=FALSE}
trip <- select(trip, -start_lat, -start_lng, -end_lat, -end_lng)
trip$started_at <- ymd_hms(trip$started_at)
trip$ended_at <- ymd_hms(trip$ended_at)
difftime <- difftime(trip$ended_at, trip$started_at, units="mins")
duration <- as.numeric(difftime)
duration <- round(duration, digits = 0)
trip <- mutate(trip, duration)
trip <- mutate(trip, day_of_week = wday(started_at, label = TRUE))
trip <- mutate(trip, month = format(started_at, "%b"))
trip$month <- ordered(trip$month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
anyNA(trip)
head(trip)

```

## Graph

```{r graph, echo=FALSE}

user <- trip %>%
        group_by(member_casual) %>%
        summarize(total = n()) %>%
        mutate (total_user = sum(total)) %>%
        group_by(member_casual) %>%
        summarize(percentage_user = total/total_user)
        user$percentage_user <- as.numeric(user$percentage_user)
pie(x = user$percentage_user,           
        labels = c("Casual 36%","Annual 64%"), 
        main = "The Percentage of Membership",
        col = blues9, 
        clockwise = TRUE,  
        init.angle = 90,  
        border = "black",  
        radius = 1)

time <- trip %>%
        group_by(member_casual) %>%
        summarize (total = sum(duration)) %>%
        mutate(total_time = sum(total)) %>%
        group_by(member_casual) %>%
        summarize(percentage_time = total/total_time)
        time$percentage_time <- as.numeric(time$percentage_time)
pie(x = time$percentage_time,           
        labels = c("Casual 56%","Annual 44%"), 
        main = "Percentage Of Duration",
        col = blues9, 
        clockwise = TRUE,  
        init.angle = 90,  
        border = "black",  
        radius = 1)

dayofweekcasual <- trip %>%
             filter(member_casual =="casual") %>%
             group_by(member_casual, day_of_week) %>% 
             summarise(numberofrides = n(),.groups = "drop")
ggplot(data = dayofweekcasual, aes(x = day_of_week, y = numberofrides)) + 
       geom_bar(stat = "identity", fill = "blue", color = "black") +
       labs(x = "Day Of Week", y = "Number Of Rides", title = "Casual Membership Daily usage")
  
dayofweekannual <- trip %>%
             filter(member_casual =="member") %>%
             group_by(member_casual, day_of_week) %>% 
             summarise(numberofrides = n(),.groups = "drop")
ggplot(data = dayofweekannual, aes(x = day_of_week, y = numberofrides)) + 
       geom_bar(stat = "identity", fill = "blue", color = "black") +
       labs(x = "Day Of Week", y = "Number Of Rides", title = "Annual Membership Daily usage")

monthlycasual <- trip %>%
           filter(member_casual =="casual") %>%
           group_by(member_casual, month) %>% 
           summarise(numberofrides = n(),.groups = "drop")
ggplot(data = monthlycasual, aes(x = month, y = numberofrides)) + 
       geom_bar(stat = "identity", fill = "blue", color = "black") +
       labs(x = "Month", y = "Number Of Rides", title = "Casual Membership Monthly Usage")

monthlyannual <- trip %>%
           filter(member_casual =="member") %>%
           group_by(member_casual, month) %>% 
           summarise(numberofrides = n(),.groups = "drop")
ggplot(data = monthlyannual, aes(x = month, y = numberofrides)) + 
        geom_bar(stat = "identity", fill = "blue", color = "black") +
        labs(x = "Month", y = "Number Of Rides", title = "Annual Membership Monthly Usage")

bikecasual <- trip %>%
        filter(member_casual =="casual") %>%
        group_by(member_casual, rideable_type) %>% 
        summarise(numberofrides = n(),.groups = "drop")
ggplot(data = bikecasual, aes(x = rideable_type, y = numberofrides)) + 
       geom_bar(stat = "identity", fill = "blue", color = "black") +
       labs(x = "Type Of Ride", y = "Nuymber Of Rides", title = "Casual Membership Bike Usage")

bikeannual <- trip %>%
       filter(member_casual =="member") %>%
       group_by(member_casual, rideable_type) %>% 
       summarise(numberofrides = n(),.groups = "drop")
ggplot(data = bikeannual, aes(x = rideable_type, y = numberofrides)) + 
       geom_bar(stat = "identity", fill = "blue", color = "black") +
       labs(x = "Type Of Ride", y = "Nuymber Of Rides", title = "Annual Membership Bike Usage")